### 2015-05-13 whoop.tex
update

The simple concurrent programming model discussed in Section~\ref{sec:symbolicpairwise} is deliberately idealistic to describe our symbolic verification technique. In practice, the Linux drivers we analyze are written in the C language, we do not know upfront which are the entry points for a driver, we do not have a named set of locks, and rather than having a given set of named shared variables, we have arbitrary memory accesses via pointers. We now explain how we have taken the conceptual ideas from Section~\ref{sec:symbolicpairwise} and used them to build \whoop, a practical tool for soundly detecting data races in drivers.

@@

\medskip\noindent\textbf{Optimizations }
%
We have implemented various optimizations to increase the precision and performance of \whoop.

In our experience, one of the most effective precision optimizations is to enrich \whoop with information regarding \emph{kernel-imposed serialization}. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized. This is an ongoing manual effort: the more drivers we study, the more such properties we discover, which we can then exploit to make \whoop more precise.

@@

A performance optimization that \whoop uses is to soundly reduce the checked shared variables. If a shared variable $s$ is accessed only by one entry point in a pair, then it is safe to not check $s$ for this pair. This reduces the burden of the SMT solver and speeds up the verification process.

>>>>

The simple concurrent programming model discussed in Section~\ref{sec:symbolicpairwise} is deliberately idealistic to describe our symbolic verification technique. In practice, the Linux drivers we analyze are written in the C language, we do not know upfront which are the entry points for a driver, we do not have a named set of locks, and rather than having a given set of named shared variables, we have arbitrary memory accesses via pointers. We now explain how we have taken the conceptual ideas from Section~\ref{sec:symbolicpairwise} and used them to build \whoop, a practical, fully automatic tool for soundly detecting data races in drivers.

@@

\medskip\noindent\textbf{Optimizations }
%
We have implemented various optimizations to increase the precision and performance of \whoop. In our experience, the following two are the most effective optimizations:

We enrich \whoop with information regarding \emph{kernel-imposed serialization}, to increase precision. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized. This is an ongoing manual effort: the more drivers we study, the more such properties we discover, which we can then exploit to make \whoop more precise.

@@

We soundly reduce the checked shared variables to increase the performance of \whoop. If a shared variable $s$ is accessed by only one entry point in a pair, then it is safe to not check $s$ for this pair. This can reduce the burden of the SMT solver and speedup the verification process.
